---
title: "Individual HW03"
author: "Your Name Here"
---

```{r}
library(bayesrules) # R package for our textbook
library(tidyverse) # Collection of packages for tidying and plotting data
library(janitor) # Helper functions like tidy and tabyl
```

# BR Exercise 5.1 a, d, f

For each situation below, tune an appropriate Gamma(s, r) prior model for $\lambda$

(a) The most common value of $\lambda$ is 4, and the mean is 7.
(d) The most common value of $\lambda$ is 14, and the variance is 6.
(f) The mean of $\lambda$ is 22, and the variance is 3.

# BR Exercise 5.7 

Let $\lambda$ be the average number of goals scored in a Women's World Cup game. We'll analyze $\lambda$ by the following Gamma-Poisson model: 
$$Y_i | \lambda \sim \text{Poisson}(\lambda)$$
$$\lambda \sim \text{Gamma}(1, 0.25)$$

(a) Plot and summarize our prior understanding of $\lambda$
(b) Why is the Poisson model a reasonable choice for $Y_i$?
(c) The `wwc_2019_matches` data in the {fivethirtyeight} package includes the number of goals scored by the two teams in each 2019 Women's World Cup match. Define, plot, and discuss the total number of goals scored per game

```{r}
library(fivethirtyeight)
data("wwc_2019_matches")
wwc_2019_matches <- wwc_2019_matches %>% 
  mutate(total_goals = score1 + score2)
```

(d) identify the posterior model of $\lambda$ and verify your answer using `summarize_gamma_poisson()`
(e) Plot the prior, likelihood, and posterior of $\lambda$. Describe the evolution in your understanding of $\lambda$ from the prior to posterior. 

# BR Exercise 5.15

Below are kernels for Normal, Poisson, Gamma, Beta, and Binomial models. Identify the appropriate model with specific parameter values.

(a) $f(\theta) \propto 0.3^\theta 0.7^{16 - \theta}$ for $\theta \in \{0, 1, 2 ..., 16\}$
(b) $f(\theta) \propto 1/\theta!$ for $\theta \in \{0, 1, ..., \infty\}$
(c) $f(\theta) \propto \theta^4(1-\theta)^7$ for $\theta \in [0,1]$
(d) $f(\theta) \propto e^{-\theta^2}$

# TBA Mon

# TBA Mon

# TBA Mon

# TBA Wed

# TBA Wed




