---
title: Evaluating Regression Models
subtitle: Day 17
title-slide-attributes:
  data-background-color: "#e2583eff"
  data-slide-number: none
format: 
  revealjs:
    incremental: false
    scrollable: false
auto-stretch: true
editor_options: 
  chunk_output_type: console
editor: 
  markdown: 
    wrap: 72
execute: 
  warning: false
  message: false
---

```{r}
#| echo: false
#| warning: false
#| message: false

library(countdown)
library(tidyverse)
library(bayesrules)
library(janitor)
library(patchwork)
library(rstan)
library(bayesplot)

theme_set(theme_minimal(base_size = 16, base_family = "Atkinson Hyperlegible"))
```

## Plan for today

- Exam recap
- Evaluating regression models

## What is the purpose of exams? 

- Provide the instructor with feedback about what concepts/topics the class as a whole understands, and which did not go as well. 
- Provide the student with feedback about what concepts/topics you understand, and which might need more time/practice
- Result in scores that summarize your performance about a (biased) selection of course material on a specific day

## 

Let $Y_i$ be the $i$th student's exam score, and assume:

$$Y_i \sim N(\mu, \sigma^2)$$

## Prior for $\mu$

Typically, I like my exams to have a mean around 86-87%, but feel they are "fair" if the mean is around 85%. In a class like Stat120 that I have taught a bunch, it is almost always between 85-89%. In classes that I don't teach as often, I'm more likely to write an exam that is too hard or too easy, and I think the mean could be as low as 65% or as high as 95%. 

$$\mu \sim N(.85, .1)$$

## Prior for $$\sigma$$

I think most students will score within 10% of the mean (ideally 75-95%), and almost all within 20% of the mean (65-100%).

```{r}
#| layout-ncol: 2


summarize_gamma(1.5, 5)
plot_gamma(1.5, 5, mean = TRUE, mode = TRUE)
```

## 

```{r}
grades_model <- "
  data {
    int<lower=1> n;
    real<lower=0, upper=1> Y[n];
  }

  parameters {
    real<lower=0, upper=1> mu;
    real<lower=0> sigma;
  }

  model {
    Y ~ normal(mu, sigma) T[0,1];
    mu ~ normal(.85, .1) T[0,1];
    sigma ~ gamma(1.5, 5);
  }
"
```



## How I interpret exam grades: {.smaller}

- **High 80s and above**: great job! You demonstrated a really solid understanding of the material. You might have made a few small mistakes or 1-2 larger errors, but I do not have major concerns about your understanding.
- **70 to mid-80s**: You probably had some high points and some weak points on the exam. I am not overly concerned about your learning, but there is room for improvement. A good final course grade is still well within your reach, but you might want to make some changes in how you're preparing for exams. 
- **Below 70**: There are either multiple key concepts that you are missing, or you had a bad exam day. To end up with a good final grade in the class, you'll want to make major changes to how you're approaching coursework and/or preparing for the next exam

## Corrections are designed to: 

- give you a chance to revisit key concepts that you missed on the exam
- mitigate the impact that a bad exam day has on your final grade

**If your score was below 80%**: 

- you can submit written corrections to earn back 30% of the points you lost, with corrected grades capped at 80% 

**If your score was below 88%**: 

- you can take a 15 minute make-up quiz on Friday of Week 7 to earn back 30% of the points you lost, with corrected grades capped at 88%



