---
title: Posterior estimation and hypothesis testing
subtitle: Day 13
title-slide-attributes:
  data-background-color: "#e2583eff"
  data-slide-number: none
format: 
  revealjs:
    incremental: false
    scrollable: false
auto-stretch: true
editor_options: 
  chunk_output_type: console
editor: 
  markdown: 
    wrap: 72
execute: 
  warning: false
  message: false
---



```{r}
#| echo: false
#| warning: false
#| message: false

library(countdown)
library(tidyverse)
library(bayesrules)
library(janitor)
library(patchwork)
library(rstan)
library(bayesplot)

theme_set(theme_minimal(base_size = 16, base_family = "Atkinson Hyperlegible"))
```

## Posterior Inference

### 1. Estimation

<br> 

<br>

### 2. Hypothesis Testing

<br> 

<br>

### 3. Prediction


## Running Example

```{r}
library(bayesrules)
library(tidyverse)
data("pulse_of_the_nation")
```

## What proportion of US adults do not believe in climate change?

```{r}
pulse_of_the_nation %>%
  count(climate_change)
```

::: callout-note
## Warm up

Let $\pi \sim$ Beta(1,2) and $Y | \pi \sim$ Binom(n, $\pi$). What is the distribution of the posterior $\pi | Y$?

:::

## 

```{r}
plot_beta_binomial(alpha = 1, beta = 2, y = 150, n = 1000)
```

## Bayesian vs Frequentist inference {.center}


In a **frequentist** analysis, $\pi$ is viewed as a fixed, unknown quantity. We make inferences about $\pi$ using data $Y$

<br>

In a **Bayesian** analysis,  $\pi$ is viewed as a random variable. We make inferences about $\pi$ using the posterior, which combines the prior and likelihood


## Example: a quick frequentist analysis

Point estimate: 150/1000

95% frequentist confidence interval: 

$$0.15 \pm 0.02 = (0.13, 0.17)$$

- [ ] There is a 95% chance that $\pi$ is between 0.13 and 0.17

<br> 

- [ ] There's a 95% chance that the survey results we happened to get produced an interval that covers $\pi$

<br>

## Example: a quick frequentist analysis

$H_0: \pi = 0.13$

$H_A: \pi > 0.13$

p-value is 0.0333

- [ ] Given our polling results, there's a 3.3% chance that $\pi$ exceeds 13%

<br> 

- [ ] If in fact $\pi$ were only 0.13, there's a 3.3% chance that we would have gotten a survey where 150 of 1000 people did not believe climate change is real



