---
title: Poisson and NB Regression
subtitle: Day 20
title-slide-attributes:
  data-background-color: "#e2583eff"
  data-slide-number: none
format: 
  revealjs:
    incremental: false
    scrollable: false
auto-stretch: true
editor_options: 
  chunk_output_type: console
editor: 
  markdown: 
    wrap: 72
execute: 
  warning: false
  message: false
---

```{r}
#| echo: false
#| warning: false
#| message: false

library(countdown)
library(tidyverse)
library(tidybayes)
library(bayesrules)
library(janitor)
library(patchwork)
library(rstan)
library(rstanarm)
library(bayesplot)
library(broom.mixed)
library(ggrepel)

theme_set(theme_minimal(base_size = 16, base_family = "Atkinson Hyperlegible"))
```

## Normal Regression

$$Y_i | \vec{X_i} \sim N(\mu_i, \sigma^2)$$
$$\mu_i = \beta_0 + \beta_1 X_1 + ... + \beta_k X_k$$

## 

::::: columns
::: {.column width="50%"}
### Usual Regression Assumptions:

:::

::: {.column width="50%"}
### Bayesian Checks:

:::
:::::

## Example 1: number of anti-discrimination laws

The `equality_index` data in the {bayesrules} package includes the observed number of anti-discrimination laws in each state,  $Y$ . Let's model  $Y$ by a state’s `percent_urban` and its `historical` presidential voting patterns. The `historical` variable has 3 levels:

- `dem` = the state typically votes for the Democratic candidate
- `gop` = the state typically votes for the Republican (GOP) candidate
- `swing` = the state often swings back and forth between Democratic and Republican candidates

## EDA

```{r}
#| echo: false 


data("equality_index")

p1 <- ggplot(equality_index, aes(x = laws)) + 
  geom_histogram(col = "white")

p2 <- ggplot(equality_index, aes(y = laws, x = percent_urban)) + 
  geom_point(size = .5)

p3 <- ggplot(equality_index, aes(y = laws, x = historical)) + 
  geom_violin(aes(fill = historical), alpha = .6) +
  geom_point(size = .5) +
  theme(legend.position = "none")

p1 / (p2 + p3)
```

## 

```{r}
equality_index |>
  filter(laws > 150)

equality <- equality_index |> filter(laws < 150) 
```

## EDA 

```{r}
#| echo: false 

p1 <- ggplot(equality, aes(x = laws)) + 
  geom_histogram(col = "white")

p2 <- ggplot(equality, aes(y = laws, x = percent_urban)) + 
  geom_point(size = .5)

p3 <- ggplot(equality, aes(y = laws, x = historical)) + 
  geom_violin(aes(fill = historical), alpha = .6) +
  geom_point(size = .5) +
  theme(legend.position = "none")

p1 / (p2 + p3)
```

## Normal model

```{r}
# Simulate the posterior Normal regression model
normal_model <- stan_glm(
  laws ~ percent_urban + historical, data = equality,
  family = gaussian,
  chains = 4, iter = 5000*2, seed = 84735, refresh = 0)
```

## What priors did I use? {.smaller}

```{r}
prior_summary(normal_model)
```

## Posterior predictive check

```{r}
pp_check(normal_model, plotfun = "hist", nreps = 15)  + 
  geom_vline(xintercept = 0, col = "darkred", linetype = "dashed")
```

## "Usual" regression check

```{r}
#| echo: false
#| fig-height: 3

p1 <- ggplot() + 
  geom_histogram(aes(x = normal_model$fitted.values), col = "white", bins = 15) + 
  labs(x = ".fitted", y = "")

p2 <- ggplot() + 
  geom_point(aes(x = normal_model$residuals, y = normal_model$fitted.values), size = .5)  + 
  labs(x = ".residuals", y = ".fitted")

p1 + p2
```

## Poisson Regression

$$Y_i | \vec{X_i} \sim \text{Pois}(\lambda_i)$$
$$\log(\lambda_i) = \beta_0 + \beta_1 X_1 + ... + \beta_k X_k$$

## Aside: GLMs

Define **data model** $$Y_i \sim F(\theta)$$

In a GLM, $$g(E(Y_i | \theta)) = \beta_0 + \beta_1 X_1 + ... + \beta_k X_k$$

$g$ is determined by the data model and called the **link function**

::: callout-note
## Question

What is the link function for normal regression? Poisson regression?
:::

## Setting up poisson model + check priors

```{r}
# Simulate the prior Poisson regression model
poisson_prior <- stan_glm(
  laws ~ percent_urban + historical, data = equality,
  family = poisson,
  prior_PD = TRUE,
  chains = 4, iter = 5000*2, seed = 84735, refresh = 0)
```

```{r}
#| output-location: slide

equality %>% 
  add_epred_draws(poisson_prior, ndraws = 50) %>%
  ggplot(aes(x = percent_urban, y = laws, color = historical)) +
    geom_line(aes(y = .epred, 
                  group = interaction(historical, .draw)), 
              alpha = .6, 
              size = .5) +
  ylim(c(0,100))
```

## Fit posterior + pp check

```{r}
poisson_model <- update(poisson_prior, prior_PD = FALSE)
```

```{r}
#| layout-ncol: 2


pp_check(poisson_model, plotfun = "hist", nreps = 5) 
pp_check(poisson_model)
```

## Interpreting posterior

```{r}
equality %>%
  add_epred_draws(poisson_model, ndraws = 50) %>%
  ggplot(aes(x = percent_urban, y = laws, color = historical)) +
    geom_line(aes(y = .epred, group = interaction(historical, .draw)), 
              alpha = .1) +
    geom_point(data = equality, size = 0.1) + 
  scale_color_viridis_d(end = .75, option = "plasma")
```

## Posterior estimates

```{r}
tidy(poisson_model, conf.int = TRUE, conf.level = 0.80)
```

::: callout-note
## Check: 

- What is the equation for the linear predictor?
- What is the expected value for $Y_i | \vec{X_i}$?
:::

## How do you interpret $\hat \beta_{\% urban} = 0.0164$

## How do you interpret $\hat \beta_{\% swing} = -0.608$

If `percent_urban` remains constant... 

(a) The number of anti-discrimination laws tends to decrease by roughly 54 percent for every extra swing state.
(b) swing states tend to have 54 percent as many anti-discrimination laws as dem leaning states.
(c) swing states tend to have 0.54 fewer anti-discrimination laws than dem leaning states.

## Posterior prediction: Minnesota

```{r}
equality %>% 
  filter(state == "minnesota")
```

## 

```{r}
# Predict number of laws for each parameter set in the chain
set.seed(84735)
as.data.frame(poisson_model) %>% 
  mutate(log_lambda = `(Intercept)` + percent_urban*73.3 + 
           historicalgop*0 + historicalswing*0,
         lambda = exp(log_lambda),
         y_new = rpois(20000, lambda = lambda)) |>
  ggplot(aes(x = y_new)) +
  stat_count() + geom_vline(xintercept = 4, col = "darkred", linetype = "dashed")
```

## Drawback of Poisson regression

```{r}
#| echo: false
#| fig-height: 3
library(ggh4x)


p1 <- pulse_of_the_nation |>
  filter(books < 200) |>
  ggplot(aes(x = books)) + 
  geom_histogram(aes(y = after_stat(density)), col = "white") + 
  stat_theodensity(distri = "pois", col = "darkorange") +
  labs( x = "x")

set.seed(102925)
p2 <- tibble(
  x = rnbinom(50, mu = 5, size = 2)
) |>
  ggplot(aes(x = x)) + 
  geom_histogram(aes(y = after_stat(density)), bins = 15, col = "white") + 
  stat_theodensity(distri = "pois", col = "darkorange")

p1 + p2

```

::: callout-note
## Overdispersion

A random variable  $Y$ is overdispersed if the observed variability in  $Y$ exceeds the variability expected by the assumed probability model of  $Y$ 
:::

## 

::: callout-note
## Negative Binomial Probabilty model

Let random variable  $Y$ be some count, $Y \in \{0, 1, ...,\}$, that can be modeled by the Negative Binomial with mean parameter  $\mu$ and reciprocal dispersion parameter $r$: 

$$Y \sim \text{NegBin}(\mu, r)$$

Then $Y$ has conditional pmf

$$f(y | \mu, r) = {y + r -1 \choose k} (\frac{r}{\mu + r})^r (\frac{\mu}{\mu+r})^y$$

and

$$E(Y|\mu, r) = \mu$$ 

$$\text{Var}(Y | \mu, r) = \mu + \frac{\mu^2}{r}$$

:::
